---
title: "SamplingStrata workflow with drake"
subtitle: "Continuous method"
author: "Giulio Barcaroli"
date: "`r Sys.Date()`"
# output: rmarkdown::html_vignette
output: 
  bookdown::html_document2:
    df_print: kable
    highlight: tango
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: yes
    toc_title: spsur vs spse
    # code_folding: hide
vignette: >
  %\VignetteIndexEntry{SamplingStrata workflow with drake}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
options(width = 999)
options(warn=-1)
knitr::opts_chunk$set(fig.width=6, fig.height=4,
  collapse = TRUE,
  comment = "#>"
)
```


# Workflow definition

First we define packages:

```{r packages, results='hide', message=FALSE, warning=FALSE}
library(drake)
library(SamplingStrata)
```

then 'swissmunicipalities' example data (with only 2 domains):

```{r load_datasets}
data(swissmunicipalities)
swissmun <- swissmunicipalities[swissmunicipalities$REG < 3,
                                c("REG","COM","Nom","HApoly",
                                  "Surfacesbois","Surfacescult",
                                  "Airbat","POPTOT")]
head(swissmun)
```

Finally, the plan:

```{r plan}
plan <- drake_plan(
  # Definition of the sampling frame
  frame = buildFrameDF(
    df = swissmun,
    id = "COM",
    domainvalue= "REG",
    X = c("HApoly","POPTOT"),
    Y =c("Surfacesbois", "Airbat")),
  # Definition of precision constraints
  cv = as.data.frame(list(
    DOM = rep("DOM1",length(unique(swissmun$REG))),
    CV1 = rep(0.10,length(unique(swissmun$REG))),
    CV2 = rep(0.10,length(unique(swissmun$REG))),
    domainvalue= c(1:length(unique(swissmun$REG))))),
  # Solution with kmeans
  kmean = KmeansSolution2(frame=frame,
                          errors=cv,
                          maxclusters=5),
  # Determination of number of strata for each domain
  nstrat = tapply(kmean$suggestions,
                  kmean$domainvalue,
                  FUN=function(x) length(unique(x))),
  # Optimisation
  solution = optimStrata(method="continuous",
                        framesamp=frame,
                        errors= cv,
                        nStrat=nstrat,
                        iter = 25,
                        pops= 10),
  # Optimal strata
  strata_structure = summaryStrata(solution$framenew,
                                   solution$aggr_strata,
                                   progress=FALSE),
  # Expected CVs with optimal strata
  expected_CV1 = expected_CV(solution$aggr_strata),
  # Adjust allocation with affordable sample size
  sample_size = 150,
  adjustedStrata = adjustSize(size=sample_size,strata=solution$aggr_strata,cens=NULL),
  # New expected CVs
  expected_CV2 = expected_CV(adjustedStrata),
  # Evaluation by simulation
  eval = evalSolution(frame = solution$framenew, 
                     outstrata = adjustedStrata, 
                     nsampl = 500,
                     progress = FALSE), 
  # Final selection of the sample
  sample = selectSample(solution$framenew,adjustedStrata)
)
```

# Execution of the workflow

We have just defined this plan:

```{r list_plan}
print(plan)
```

visualised in this way:

```{r graph}
vis_drake_graph(plan, ncol_legend = NULL,
                main="Optimization with 'continuous' method")
```

and now we execute:

```{r execution}
make(plan)
```


# Output inspection

First of all, the total size of the optimized sample:

```{r }
options(width = 999)
loadd(solution)
sum(solution$aggr_strata$SOLUZ)
```

Then, the structure of the optimized strata:

```{r }
loadd(strata_structure)
strata_structure

```

The expected CVs given the optimized strata:

```{r }
loadd(expected_CV1)
expected_CV1
```

and after the adjustment of the allocation in the strata based on the affordable sample size:

```{r }
loadd(expected_CV2)
expected_CV2
```

The results of simulation with the adjusted allocation:

```{r }
loadd(eval)
eval$coeff_var
eval$rel_bias
```

Finally, the selection of the sample from the adjusted strata:

```{r }
loadd(sample)
head(sample)
```


